\documentclass[9pt,aspectratio=169,handout]{beamer}

\usepackage{scalerel}

\usetheme{graham}

\title{Quadratic Equations solutions}
\subtitle[Graham Middle School]{Graham Middle School Math Olympiad Team}

\newcommand\Mydiv[2]{%
$\strut#1$\kern.25em\smash{\raise.3ex\hbox{$\big)$}}$\mkern-8mu
        \overline{\enspace\strut#2}$}

\setcounter{MaxMatrixCols}{20}
\newcommand{\Mod}[1]{\ (\mathrm{mod}\ #1)}
\newcommand{\longdiv}{\smash{\mkern-0.43mu\vstretch{1.31}{\hstretch{.7}{)}}\mkern-5.2mu\vstretch{1.31}{\hstretch{.7}{)}}}}

\DeclareMathOperator{\lcm}{lcm}

\usepackage{luamplib}
\mplibsetformat{metafun}
\mplibtextextlabel{enable}
\everymplib{input repere; input macros; beginfig(1);}
\everyendmplib{endfig;}

\begin{document}
\maketitle


\begin{frame}{Exercises 1-4}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{E1.} Use factoring to find the roots of 
        
        $x^2 – 22x – 48 = 0$.
      \end{problem}
      Let $x^2 - 22x - 48 = (x - a)(x - b)$, so $a + b = 22$ and $ab = -48$. Looking into divisors of $-48$ we may find that $24$ and $-2$ has a sum of $22$, so the root of the equation are $\boxed{-2}$ and $\boxed{24}$. 

      \begin{problem}
        \textbf{E2.} Complete the square to find the possible values of $x$ for which $x^2 + 4x + 3 = 0$?
      \end{problem}
      To complete square for $x^2 + 4x$ we need to add $4$, since $(x+2)^2 = x^2 + 4x + 4$. So, in our equation we have $x^2 + 4x + 3 + 1 = 1$ or $(x+2)^2 = 1$. This gives us $x + 2 = \pm 1$ and $x = \boxed{-3}$ or $\boxed{-1}$.

    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{E3.} What is the value of $i^3$?
      \end{problem}
      $i^3 = (i)^2 \times i = -1 \times i = -i$.
      \begin{problem}
        \textbf{E4.} For what value(s) of $x$ does the fraction of $3$ raised to the power $x^2$ over $3$ raised to the power $3x$ equal one-ninth?  
        
        (\emph{Hint}: If all exponents have the same base, then we can solve the problem by equating the exponents.
      \end{problem}
      We need to solve this equation:
      \[ \frac{3^{x^2}}{3^{3x}} = \frac{1}{9} = \frac{1}{3^2}.\]
      Which can be rewritten as $3^{x^2} \times 3^{-3x} = 3^{-2}$, or, using our hint, $x^2 - 3x = -2$, $x^2 - 3x + 2 = (x - 1)(x - 2) = 0$ and our equation has two roots $x = \boxed{1}$ and $\boxed{2}$. Plugging them back into our original equation we can check that they are works.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Exercises 5-8}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{E5.} Find the roots of $x = \dfrac{28}{x - 3}$.
      \end{problem}
      $x(x - 3) = 28$ or $x^2 - 3x - 28 = (x - 7)(x + 4) = 0$ and two roots $x = \boxed{-4}$ and $\boxed{7}$. Plugging them back into our original equation we can check that they are works.
      \begin{problem}
        \textbf{E6.} If $b$ and $c$ are both rational numbers and one of the roots of $x^2 + bx + c = 0$ is $3 + \sqrt{2}$, find $b$ and~$c$.
      \end{problem}
      The roots are $\dfrac{-b}{2} \pm \sqrt{\dfrac{b^2 - 4c}{4}}$, so we may see that rational and irrational parts of both roots should be the same and the second root is $3 - \sqrt{2}$. $x^2 + bx + c = (x - (3 +\sqrt{2}))(x - (3 - \sqrt{2})) = x^2 - 6x + 7 = 0$. So $\boxed{b = -6}$ and $\boxed{c = 7}$.
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{E7.} For how many different integer values of $b$ are both roots of $x^2 + bx - 16 = 0$ integers?
      \end{problem}
      Let $x^2 + bx - 16 = (x + p)(x + q) = 0$, so $pq = -16$ and $b = p + q$. Since $-16 = 1\cdot -16 = 2\cdot -8 = 4\cdot -4 = 8\cdot -2 = 16\cdot -1 = -1\cdot 16= -2\cdot 8= -4\cdot 4= -8\cdot 2= -16\cdot 1$. We got $\boxed{5}$ different values for $b$: $-15$, $-6$, $0$, $6$, and $15$.
      \begin{problem}
        \textbf{E8.} Let $m$ and $n$ be roots of: $x^2 - 60x + 864 = 0$.  Find a polynomial with roots $m + 1$ and $n + 1$.
      \end{problem}
      $(x - (m + 1))(x - (n +1)) = x^2 - (m + n + 2)x + (mn + m + n + 1) = 0$ has roots $m +1$ and $n +1$, since $m + n = 60$ and $mn = 864$, we got $\boxed{x^2 - 62x + 925 = 0}$.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Challenge problems 1 - 2}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{CP1.} Find the minimum possible value of the absolute value of $(m - n)$, where $m$ and $n$ are integers satisfying $m + n = mn - 2021$. 
        
        (\emph{Hint}: could completing the square be useful here if the variables were all grouped on one side of the equation?)
      \end{problem}
      $mn - m - n - 2021 = (m - 1)(n - 1) - 2022$ or $(m - 1)(n - 1) = 2022$, for $|m - n|$ to be as minimal as possible we need $m$ and $n$ as close as possible. The closest factors of $2022$ are $6$ and $337$, so answer is $337 - 6 = \boxed{331}$.
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{CP2.} (For fun) In the novel, “The Curious Incident of the Dog in the Nighttime,” a student in England taking his A-level college entrance exam in maths was given the following question: 
        
        Prove that a triangle with sides that can written in the form $n^2 +1$, $n^2 -1$ and $2n$ (where $n > 1$) is right-angled.
      \end{problem}
      Since $(n^2 + 1)^2 = n^4 + 2n^2 + 1 = n^4 - 2n^2 + 1 + 4n^2 = (n^2 - 1)^2 + (2n)^2$, and using \emph{Converse of Pythagoras Theorem} we got that triangle with sides  $n^2 +1$, $n^2 -1$ and $2n$ (where $n > 1$) is right-angled.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Challenge problems 3 - 4}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{CP3.} Let $m$ and $n$ be roots of the polynomial $x^2 - 60x + 899 = 0.$  What is $m^2 + n^2$? 
        
        (\emph{Hint}: think about how $m^2 + n^2$ can be rewritten in terms of the sum and product of the roots $m$ and $n$).
      \end{problem}

      $m^2 + n^2 = m^2 + 2mn + n^2 - 2mn = (m + n)^2 - 2mn = (60)^2 - 2\cdot 899 = 3600 - 1798 = \boxed{1802}$.
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{problem}
        \textbf{CP4.} Find all real values of $n$ such that $2^{2n} + 2^n + 1 = 73$.  
        
        (\emph{Hint}:  What substitution would turn this into a quadratic?)
      \end{problem}
      Let $y = 2^n$, then our equation turns into $y^2 + y + 1 = 73$ or $y^2 + y - 72 = (y - 8)(y + 9) = 0$. So $y = 8$ or $-9$. If $2^n = 8$, $n = \boxed{3}$, if $2^n = -9$ we don't have solutions since $2^n > 0$ for any real $n$.
    \end{column}
  \end{columns}
\end{frame}

% \begin{frame}{Title}
%   \begin{columns}[T]
%     \begin{column}{0.5\textwidth}
%     \end{column}
%     \begin{column}{0.5\textwidth}
%     \end{column}
%   \end{columns}
% \end{frame}

\end{document}